---
title: "TidyTuesday: Age gaps in Hollywood movies."
author: "Hrólfur Ásmundsson"
date: "2023-02-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,warning=FALSE,message=FALSE}
library(tidyverse)
library(tidytuesdayR)
```

This weeks data set comes from [hollywoodagegap.com](hollywoodagegap.com). As usual it is easy very to load the data using the **`tidytuesdayR`** package

```{r,cache=TRUE,eval=FALSE}
tuesdata <- tidytuesdayR::tt_load(2023, week = 7)
# save and set eval=FALSE so I don't need to call github any time I Knit :)
write_csv(tuesdata$age_gap,"age_gap.csv")  
```

Let's begin by exploring the data set

```{r,message=FALSE,warning=FALSE}
data <- read_csv("age_gap.csv")
head(data)
```

We have a lot of different data that we could look at and many interesting things to learn from it. I think I find it most interesting to see if the age difference between women and men is getting smaller in modern times.

## Simple scatter plot To see this lets first simply make a scatter plot of age difference as a function of the release year.

```{r,message=FALSE,warning=FALSE}
g1<- data %>%
  ggplot(aes(x=release_year,y=age_difference))+
  geom_point()
g1
```

From this we see that points get more dense at lower age difference after the turn of the century, but it could also be because there are a lot more movies made. Thankfully not all the couples in movies are heterosexual. In fact there are in total 23 different same sex couples. Let's also identify them on the figure.

```{r,message=FALSE,warning=FALSE}
n_same_sex <- sum(data$character_1_gender==data$character_2_gender)
n_same_sex

g2<- data %>%
  mutate(samesex=data$character_1_gender==data$character_2_gender) %>%
  ggplot(aes(x=release_year,y=age_difference,shape=samesex))+
  geom_point()
g2
```

Our figure does not yet distinguish between when the woman or the man is older so lets introduce a color scheme for that. We can make use of the fact that the data set has been made so that `charecter_1` is always older than `character_2.`

```{r,message=FALSE,warning=FALSE}

g3<- data %>%
  mutate(samesex=character_1_gender==character_2_gender) %>%
  ggplot(aes(x=release_year,y=age_difference,shape=samesex,color=character_1_gender))+
  geom_point(alpha=0.8)+
  theme_minimal()+
  scale_x_discrete(name ="Release Year", 
                   limits=seq(1940,2023,10))+
  scale_y_discrete(name ="Age difference in years", 
                   limits=seq(0,50,10))+
  scale_fill_manual(values=c("man"="blue","woman"="red"),labels=c("The man","The woman"))+
  labs(title="Hollywood Age Gap",color="Who is Older?",shape="Same sex couple")

g3
```

The points are rather clustered in the lower right corner so it is not very easy to make out. So lets plot the age difference relative to the age of the woman.

```{r,warning=FALSE,message=FALSE}
g4 <- data %>%
  mutate(samesex=data$character_1_gender==data$character_2_gender) %>%
  mutate(relative_gap = if_else(character_1_gender=="man",age_difference,-age_difference)) %>% 
  ggplot(aes(x=release_year,y=relative_gap,color=character_1_gender,shape=samesex))+
  geom_point(alpha=0.5)+
  theme_minimal()+
  scale_x_discrete(name ="Release Year", 
                   limits=seq(1940,2023,10))+
  scale_y_discrete(name ="Age difference in years", 
                   limits=seq(-50,50,10),labels=abs(seq(-50,50,10)))+
  scale_color_manual(values=c("man"="blue","woman"="red"),labels=c("The man","The woman"))+
  labs(title="Hollywood Age Gap",color="Who is Older?",shape="Same sex couple",caption="Tidytuesday 2023-02-14: Data from hollywoodagegap.com")

g4
```

This is our final version of this figure, it is pretty clear, from the figure, that it is still much likelier that the man is older.

## Colored Ribbon graph To investigate this further, let's consider the mean age of women and man actors every year of the data set.

```{r,message=FALSE,warning=FALSE}
g5<- tibble(movie = c(data$movie_name,data$movie_name),
            year = c(data$release_year,data$release_year),
            age = c(data$actor_1_age,data$actor_2_age),
            gender = c(data$character_1_gender,data$character_2_gender)) %>% 
  group_by(year,gender) %>% 
  summarise(mean_age = mean(age)) %>% 
  ggplot(aes(x=year,y=mean_age,color=gender)) +
  geom_line()
g5
```

The mean age of man actors is consistently above the women, except for the year 1971, but that is due to a single movie *Harold and Maude* where the woman is 52 years older than the man.

```{r}
data %>% filter(release_year==1971) %>% select(movie_name,age_difference,actor_1_name,actor_2_name)
```
Lets visualize the difference with a colored ribbon,
```{r,message=FALSE,warning=FALSE}

g6<- tibble(movie = c(data$movie_name,data$movie_name),
            year = c(data$release_year,data$release_year),
            age = c(data$actor_1_age,data$actor_2_age),
            gender = c(data$character_1_gender,data$character_2_gender)) %>% 
  group_by(year,gender) %>% 
  summarise(mean_age = mean(age)) %>% 
  pivot_wider(names_from = gender, values_from = mean_age) %>% 
  mutate(higer = ifelse(man>woman,"man","woman")) %>% 
  ggplot(alpha=0.5) +
  geom_ribbon(aes(x=year,ymin=woman,ymax=pmin(man,woman),fill="Woman"))+
  geom_line(aes(x=year,y=man),color="blue",width=2)+
  geom_line(aes(x=year,y=woman),color="red",width=2)+
  geom_ribbon(aes(x=year,ymin=man,ymax=pmin(man,woman),fill="Man"))+
  theme_minimal()+
  scale_x_discrete(name ="Release Year", 
                   limits=seq(1940,2023,10))+
  scale_y_discrete(name ="Mean Age", 
                   limits=seq(-50,50,10),labels=abs(seq(-50,50,10)))+
  scale_fill_manual(values=c("Man"="lightblue","Woman"="pink"),labels=c("Men","Women"))+
  labs(title="Hollywood Age Gap",fill="Which are Older on Average?",shape="Same sex couple",caption="Tidytuesday 2023-02-14: Data from hollywoodagegap.com")
g6
```

## Age distributions
As our final figure we shall find the empirical distribution function for the age of the actors of each gender.
```{r,message=FALSE,warning=FALSE}
Emperical_DF <- tibble(movie = c(data$movie_name,data$movie_name),
            year = c(data$release_year,data$release_year),
            age = c(data$actor_1_age,data$actor_2_age),
            gender = c(data$character_1_gender,data$character_2_gender)) %>% 
  group_by(gender) %>% summarise(edf = list(ecdf(age)))

g7 <- tibble(age = rep(17:90,2),
                     p = c(Emperical_DF$edf[1][[1]](17:90),Emperical_DF$edf[2][[1]](17:90)),
                     gender = rep(c("man","woman"),each=74)) %>% 
  ggplot(aes(x=age,y=p,color=gender))+
  geom_line(size=1)+
  scale_color_manual(values=c("man"="blue","woman"="red"))+
  theme_minimal()+
  scale_x_discrete(name ="Actors age", 
                   limits=seq(15,90,5))+
  labs(title="Hollywood Age Gap",fill="Which are Older on Average?",shape="Same sex couple",caption="Tidytuesday 2023-02-14: Data from hollywoodagegap.com")+
  ylab(expression(P(x < age)))
g7
```
We clearly see that the men are older and last longer as possible "love interests", we see that for example there is roughly 50% chance that an actress is younger than 30 years old, while there is only 20% chance for men.

```{r}
Emperical_DF$edf[1][[1]](30)
Emperical_DF$edf[2][[1]](30)
```

